\begin{abstract}
Concurrency control is an integral component in achieving high performance in many-core databases. 
Implementing serializable transaction processing efficiently is challenging. One approach, 
\emph{serialization graph testing} (SGT) 
faithfully implements the \emph{conflict graph theorem} by aborting only those transactions 
% directly utilizes the \emph{conflict graph theorem}, which states that a 
% schedule is conflict serializable if and only if its associated conflict graph is acyclic. By aborting 
% only transactions 
that would actually
% truly 
violate serializability (introduce a cycle), thus maintaining
the required acyclic invariant. 
Alternative approaches, such as \emph{two-phase locking}, 
disallow certain valid 
% reduce the amount of accepted 
schedules to increase throughput, whereas SGT has the theoretically optimal property of accepting all 
and only
conflict serializable schedules. Historically, SGT was deemed 
% nonviable due to concerns over the 
unviable in practice due to the high
% compuatational costs of...graph.
computational costs of maintaining an acyclic graph. 
Research has however overturned this historical view by utilising the increased computational 
power available due to modern hardware.  
% this conventional wisdom has been refuted by recent research. 
% Building on this result, this paper demonstrates how graph-based concurrency control can be 
% extended to support transactions executing at weaker isolation levels, whilst maintaining the desirable 
% property of minimizing aborts. 
Furthermore, a survey of 24 databases suggests that not all transactions demand conflict serializability but different 
transactions can perfectly settle for different, weaker isolation levels which typically require relatively lower overheads.  
Thus, in such a mixed environment, providing only the isolation level required of each transaction should, in theory,  
increase throughput and reduce aborts. The aim of this paper is to extend SGT for mixed environments subject to Adya's 
\emph{mixing-correct theorem} and demonstrate the resulting performance improvement. We augment the YCSB benchmark to 
generate transactions with different isolation requirements. For certain workloads, \emph{mixed serialization graph testing} can achieve up to a 28\% increase in throughput and a 19\% 
decrease in aborts over SGT.
% \emph{Mixed serialization graph testing} (MSGT) leverages Adya's weak 
% isolation theory accepting all valid schedules under the \emph{mixing-correct theorem}. MSGT's practical 
% usefulness is illustrated by surveying isolation levels supported by 24 databases. To evaluate MSGT's 
% performance, the YCSB benchmark is augmented to generate transactions with different isolation 
% requirements. MSGT can achieve up to a 17\% increase in throughput and a 9.5\% decrease in aborts over SGT.





\keywords{Databases \and Concurrency Control \and Weak Isolation \and Serialization Graph Testing \and Mixing-Correct Theorem 
\and YCSB}
\end{abstract}