\bibliographystyle{splncs04}
\bibliography{mybibliography}
%
\begin{thebibliography}{8}
%
\bibitem{DBLP:books/aw/BernsteinHG87}
Bernstein, P., Hadzilacos, V., Goodman, N.:
Concurrency Control and Recovery in Database Systems. (1987)
%
\bibitem{DBLP:journals/cacm/EswarranGLT76}
Eswaran, K., Gray, J., Lorie, R., Traiger, I.:
The Notions of Consistency and Predicate Locks in a Database System.
Commun. {ACM}, vol. 19(11), pp. 624--63 (1976)
%
\bibitem{DBLP:journals/tods/KungR81}
Kung, H., Robinson, J.:
On Optimistic Methods for Concurrency Control.
{ACM} Trans. Database Syst., vol. 6(2), pp. 213--226 (1981)
%
\bibitem{DBLP:conf/icde/Durner019}
Durner, D., Neumann, T.: No False Negatives: Accepting All Useful Schedules in a Fast Serializable Many-Core System. 35th {IEEE} International Conference on Data Engineering, 734--745 (2019) 
%
% \bibitem{DBLP:conf/icde/AdyaLO00}
% Adya, A., Liskov, B., O'Neil, P.: Generalized Isolation Level Definitions. Proceedings of the 16th International Conference on Data Engineering, 67--78 (2000)
%
\bibitem{adya1999weak}
Adya, A.: Weak {C}onsistency: {A} {G}eneralized {T}heory and {O}ptimistic {I}mplementations for {D}istributed {T}ransactions. PhD Thesis, Massachusetts Institute of Technology (1999)
%
\bibitem{DBLP:conf/ds/GrayLPT76}
Gray, J., Lorie, R., Putzolu, G.,Traiger, I.:
{G}ranularity of {L}ocks and {D}egrees of {C}onsistency in a {S}hared {D}ata {B}ase. 
Modelling in Data Base Management Systems, 365--394 (1976)
%
\bibitem{DBLP:conf/cloud/CooperSTRS10}
Cooper, B., Silberstein, A., Tam E., Ramakrishnan, R., Sears, R.:
Benchmarking cloud serving systems with {YCSB}.
Proceedings of the 1st {ACM} Symposium on Cloud Computing, pp. 143--154 (2010)
%
% \bibitem{DBLP:conf/sosp/TuZKLM13}
% Tu, D., Zheng, W., Kohler, E., Liskov, B., Madden, S.:
% Speedy transactions in multicore in-memory databases.
% {ACM} {SIGOPS} 24th Symposium on Operating Systems Principles, 18--32 (2013)
%
% \bibitem{DBLP:conf/sigmod/0001MK15}
% Neumann, T., M{\"{u}}hlbauer, T., Kemper, A.:
% Fast Serializable Multi-Version Concurrency Control for Main-Memory Database Systems.
% Proceedings of the 2015 {ACM} {SIGMOD} International Conference on Management of Data, 677--689 (2015)
%
% \bibitem{DBLP:conf/sigmod/YuPSD16}
% Yu, X., Pavlo, A., S{\'{a}}nchez, D., Devadas, S.:
% TicToc: Time Traveling Optimistic Concurrency Control.
% Proceedings of the 2016 {ACM} {SIGMOD} International Conference on Management of Data, 1629--1642, (2016)
%
\bibitem{DBLP:journals/sigmod/PavloA16}
Pavlo, A., Aslett, M.: 
What's Really New with NewSQL?.
{SIGMOD} Rec., vol. 45(2), pp. 45--55 (2016)
%
% \bibitem{DBLP:journals/debu/StonebrakerW13}
% Stonebraker, M., Weisberg, A.,:
% The VoltDB Main Memory {DBMS}.
% {IEEE} Data Eng. Bull., vol. 36(2), pp. 21--27 (2013)
%
\bibitem{DBLP:conf/sigmod/TaftSMVLGNWBPBR20}
Taft, R. et al.:
CockroachDB: The Resilient Geo-Distributed SQL Database.
Proceedings of the 2020 ACM SIGMOD International Conference on Management of Data., 
pp. 1493--1509 (2020)
%
% \bibitem{DBLP:journals/tocs/CorbettDEFFFGGHHHKKLLMMNQRRSSTWW13}
% Corbett, J. et al.:
% Spanner: Google's Globally Distributed Database.
% {ACM} Trans. Comput. Syst.,
% vol. 31(3), pp. 1--22 (2013)
%
\bibitem{tidb}
TiDB Transaction Isolation Levels, 
\url{https://docs.pingcap.com/tidb/dev/transaction-isolation-levels}. 
Last accessed 9 May 2022
%
\bibitem{DBLP:journals/tods/FeketeLOOS05}
Fekete, A., Liarokapis, D., O'Neil, E., O'Neil, P., Shasha, D.:
Making snapshot isolation serializable.
{ACM} Trans. Database Syst.,
vol. 30(2),
pp. 492--528
(2005)
%
\bibitem{DBLP:phd/ethos/Fraser04}
Fraser, K.: Practical lock-freedom. PhD Thesis, University of Cambridge (2004)
% 
\bibitem{tpcx-iot}
Poess, M., Nambiar, R., Kulkarni, K., Narasimhadevara, C., Rabl, T., Jacobsen, H.A.:
Analysis of TPCx-IoT: The First Industry Standard Benchmark for IoT Gateway Systems. 34th {IEEE} International Conference on Data Engineering, 1519--1530 (2018) 
%
\bibitem{tpcc}
{TPC Benchmark C}, revision 5.11, 
\url{http://www.tpc.org/tpc_documents_current_versions/pdf/tpc-c_v5.11.0.pdf}. 
Last accessed 19 July 2022
%
% \bibitem{DBLP:journals/pvldb/AlvaroK20}
% Alvaro, P., Kingsbury, K.:
% Elle: Inferring Isolation Anomalies from Experimental Observations.
% Proc. {VLDB} Endow.,
% vol. 14(3),
% pp. 268--280
% (2020)
%
% \bibitem{DBLP:journals/pvldb/GanRRB020}
% Gan, Y., Ren X., Ripberger, D., Blanas, S., Wang, Y.:
% IsoDiff: Debugging Anomalies Caused by Weak Isolation.
% Proc. {VLDB} Endow.,
% vol. 13(11),
% pp. 2773--2786
% (2020)
%
% \bibitem{DBLP:conf/icde/RevilakOO11}
% Revilak, S., O'Neil, P., O'Neil, E.:
% Precisely Serializable Snapshot Isolation {(PSSI)}.
% Proceedings of the 27th International Conference on Data Engineering.
% pp. 482--493
% (2011)  
%
\end{thebibliography}